---
title: "Session Management: List, Read, and Mutate Sessions"
description: >
  How to discover, read, rename, tag, fork, and delete Claude Code sessions
  stored on disk. Covers the full session management API for building
  history browsers and audit tooling (e.g. Shiny UIs).
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Session Management: List, Read, and Mutate Sessions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

## Overview

Claude Code persists every conversation as a JSONL file under
`~/.claude/projects/<sanitized-cwd>/<session-uuid>.jsonl`.
ClaudeAgentSDK provides a pure-R API for reading and mutating these files
**without** requiring a live CLI connection — all functions operate directly
on disk.

| Function | Purpose |
|---|---|
| `list_sessions()` | Discover sessions (all projects or one project) |
| `get_session_info()` | Metadata for a single session |
| `get_session_messages()` | Full conversation transcript |
| `rename_session()` | Set a human-readable title |
| `tag_session()` | Attach a tag string (or clear it) |
| `fork_session()` | Copy session to a new UUID |
| `delete_session()` | Remove the JSONL file |

---

## Listing sessions

### All projects

```{r list-all}
library(ClaudeAgentSDK)

# Returns SDKSessionInfo objects sorted by last_modified descending
sessions <- list_sessions(limit = 20L)
length(sessions)

# Every session carries: session_id, summary, last_modified, cwd,
# first_prompt, custom_title, tag, git_branch, created_at, file_size
s <- sessions[[1]]
cat(s$session_id, "|", s$summary %||% "(no summary)", "\n")
cat("cwd:", s$cwd, "\n")
cat("first_prompt:", substr(s$first_prompt %||% "", 1, 80), "\n")
```

### One project only

```{r list-project}
# Pass the project's working directory; SDK hashes it to find the
# matching ~/.claude/projects/<hash>/ sub-directory automatically.
sessions <- list_sessions(directory = getwd(), limit = 10L)
cat("Sessions in this project:", length(sessions), "\n")
```

### Pagination

```{r paginate}
page1 <- list_sessions(limit = 10L, offset = 0L)
page2 <- list_sessions(limit = 10L, offset = 10L)
```

---

## Reading session metadata

```{r get-info}
session_id <- sessions[[1]]$session_id

info <- get_session_info(session_id)
if (!is.null(info)) {
  cat("summary:      ", info$summary %||% "(none)", "\n")
  cat("custom_title: ", info$custom_title %||% "(none)", "\n")
  cat("tag:          ", info$tag %||% "(none)", "\n")
  cat("last_modified:", format(info$last_modified), "\n")
}
```

---

## Reading conversation history

`get_session_messages()` reconstructs the conversation chain via
`parentUuid` links and returns only visible user/assistant turns in
chronological order.

```{r get-messages}
msgs <- get_session_messages(session_id)
cat("Total turns:", length(msgs), "\n\n")

for (m in msgs) {
  if (m$type == "user") {
    raw  <- m$message
    text <- if (is.character(raw$content)) raw$content else "(complex content)"
    cat("User:     ", text, "\n")
  } else {
    raw <- m$message
    if (is.list(raw$content)) {
      for (blk in raw$content)
        if (identical(blk[["type"]], "text")) cat("Assistant:", blk[["text"]], "\n")
    } else if (is.character(raw$content)) {
      cat("Assistant:", raw$content, "\n")
    }
  }
}
```

Each `SessionMessage` object has:

| Field | Type | Description |
|---|---|---|
| `type` | `"user"` or `"assistant"` | Speaker role |
| `uuid` | character | Message UUID |
| `session_id` | character | Parent session UUID |
| `message` | list | Raw parsed JSON (contains `content`, `role`, etc.) |

---

## Session mutations

All mutation functions operate directly on the JSONL files —
**no CLI connection required**.

### Rename (set custom title)

```{r rename}
rename_session(session_id, title = "Sprint 12 planning")

# Verify
info <- get_session_info(session_id)
cat("custom_title:", info$custom_title, "\n")
```

Rename uses append-only semantics: a new `title` entry is appended to the
JSONL and the most-recent-wins rule applies. The original messages are
untouched.

### Tag

```{r tag}
tag_session(session_id, tag = "reviewed")

# Clear a tag
tag_session(session_id, tag = NULL)
```

### Fork

`fork_session()` copies the transcript to a new UUID, remapping all
internal UUIDs so the fork is fully independent.

```{r fork}
forked_id <- fork_session(
  session_id,
  title = "Experiment branch",
  # Optionally slice the transcript up to a specific message:
  # up_to_message_id = "some-uuid"
)
cat("Forked:", forked_id, "\n")
```

### Delete

```{r delete}
# Permanently removes the .jsonl file
delete_session(forked_id)

# Confirm
stopifnot(is.null(get_session_info(forked_id)))
```

---

## Shiny history browser pattern

The session listing API is designed for building read-only history UIs.
A minimal Shiny app structure:

```{r shiny-pattern}
library(shiny)
library(ClaudeAgentSDK)

ui <- fluidPage(
  sidebarLayout(
    sidebarPanel(
      selectInput("session", "Session", choices = NULL),
      actionButton("load", "Load transcript")
    ),
    mainPanel(
      uiOutput("transcript")
    )
  )
)

server <- function(input, output, session) {
  # Populate selector on startup
  observe({
    sessions <- list_sessions(limit = 50L)
    labels   <- vapply(sessions, function(s) {
      paste0(
        format(s$last_modified, "%Y-%m-%d %H:%M"),
        "  ",
        s$custom_title %||% s$summary %||% s$session_id
      )
    }, character(1))
    ids      <- vapply(sessions, function(s) s$session_id, character(1))
    updateSelectInput(session, "session", choices = setNames(ids, labels))
  })

  transcript <- eventReactive(input$load, {
    req(input$session)
    get_session_messages(input$session)
  })

  output$transcript <- renderUI({
    msgs <- transcript()
    items <- lapply(msgs, function(m) {
      if (m$type == "user") {
        raw  <- m$message
        text <- if (is.character(raw$content)) raw$content else "(complex)"
        div(class = "user-turn",   strong("User: "),      text)
      } else {
        raw <- m$message
        txt <- if (is.list(raw$content)) {
          blks <- Filter(function(b) identical(b[["type"]], "text"), raw$content)
          paste(vapply(blks, `[[`, character(1), "text"), collapse = "\n")
        } else if (is.character(raw$content)) raw$content else "(complex)"
        div(class = "asst-turn", strong("Assistant: "), txt)
      }
    })
    do.call(tagList, items)
  })
}

shinyApp(ui, server)
```

---

## Notes

- **No CLI required.** All functions read/write `~/.claude/` directly.
- **`directory` parameter.** The SDK applies the same path-sanitization
  hash as the Claude Code CLI, so `directory = getwd()` always resolves to
  the correct project sub-directory.
- **Worktree support.** `list_sessions(include_worktrees = TRUE)` (the
  default) also scans git worktree paths associated with the project.
- **Mutation semantics.** `rename_session()` and `tag_session()` are
  append-only; the JSONL is never rewritten, so the originals are always
  recoverable. `delete_session()` and `fork_session()` are the only
  file-level operations.
